\LaTeXe\ guide for the Journal of Functional Programming

### 2019-04-26 papers/jfp/8-engineering.tex
Finish S8.7 (Polymorphism and multiple-output build tasks)

\subsection{Polymorphism}\label{sec-polymorphism}

@@

A more compelling example is build tasks that produce multiple output
keys~--~for example, \cmd{ghc Foo.hs} produces both \cmd{Foo.hi} and \cmd{Foo.o}.
That can be represented by having a key whose value is a pair of file names, and
whose result is a pair of file contents. From that, the task for \cmd{Foo.hi}
can be the first component of the result of the pair. Again, such an operation
can be encoded without polymorphic keys provided the pair of files (or a dummy
file representing the pair) is marked as changed if either of the contained
files change. Once again, polymorphic dependencies provide convenience rather
than power.

% \vspace{1mm}
% \begin{minted}[xleftmargin=10pt]{haskell}
% type Partition k = k -> [k]

% multi :: Eq k => Partition k -> Tasks Applicative [k] [v] -> Tasks Applicative [k] [v]
% multi partition tasks keys
%     | k:_ <- keys, partition k == keys = tasks keys
%     | otherwise = Just $ Task $ \fetch ->
%         sequenceA [ select k <$> fetch (partition k) | k <- keys ]
%   where
%     select k = fromMaybe (error msg) . lookup k . zip (partition k)
%     msg = "Partition invariants violated"
% \end{minted}
% \vspace{1mm}

>>>>

\subsection{Polymorphism and Multiple-Output Build Tasks}\label{sec-polymorphism}

@@

A more compelling example is build tasks that produce multiple output
keys~--~for example, \cmd{ghc A.hs} produces both \cmd{A.hi} and \cmd{A.o}.
That can be represented by having a key whose value is a pair of file names, and
whose result is a pair of file contents. From that, the task for \cmd{A.hi}
can be the first component of the result of the pair. Again, such an operation
can be encoded without polymorphic keys provided the pair of files (or a dummy
file representing the pair) is marked as changed if either of the contained
files change. Once again, polymorphic dependencies provide convenience rather
than power.

Note that multiple-output build tasks require one more convenience mechanism,
which is orthogonal to polymorphism. Multiple-output task descriptions partition
the set of keys into \emph{elementary subsets} e.g.
$\{\cmd{A.hi},\cmd{A.o}\} \rightarrow \cmd{ghcTask}$,
$\{\cmd{B.o}\} \rightarrow \cmd{gccTask}$, etc. It is very convenient to
allow to fetch arbitrary sets of keys, such as $\{\cmd{A.o},\cmd{B.o}\}$,
without manually decomposing them into appropriate elementary subsets. This
requires a method to efficiently construct task descriptions operating on
arbitrary sets of keys rather on elementary subsets. One way to implement such
``multiple-output task closure'' is shown below.

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
type Partition k = k -> [k]

multi :: Eq k => Partition k -> Tasks Applicative [k] [v]
                             -> Tasks Applicative [k] [v]
multi partition tasks keys
    | k:_ <- keys, partition k == keys = tasks keys
    | otherwise = Just $ Task $ \fetch ->
        sequenceA [ select k <$> fetch (partition k) | k <- keys ]
  where
    select k = fromMaybe (error msg) . lookup k . zip (partition k)
    msg = "Partition invariants violated"
\end{minted}
\vspace{1mm}

\noindent
The function \hs{multi} takes a task description \hs{tasks} that operates on
elementary key subsets, and computes its closure supporting all possible
combinations of keys. The first argument \hs{partition} specifies the elementary
key subsets supported by the original \hs{tasks}. There are two cases: (i) if a
given set of \hs{keys} is an elementary subset, we pass the query to \hs{tasks};
(ii) otherwise, we examine every key \hs{k} in \hs{keys}, \hs{fetch} its
corresponding elementary subset, and \hs{select} the appropriate value from the
result. Returning to the above example, when given the query
$\{\cmd{A.o},\cmd{B.o}\}$, this function will \hs{fetch} elementary subsets
$\{\cmd{A.hi},\cmd{A.o}\}$ and $\{\cmd{B.o}\}$, thereby depending on the
corresponding build tasks \cmd{ghcTask} and \cmd{gccTask}.

We chose not to use polymorphic tasks in the above implementation for
simplicity, but nothing prevents us from combining polymorphism with
multiple-output task closure.
